---
title: "ALY6160_GroupProject"
author: "Catherine Richard, Shivangi Vashi"
date: "2/6/2021"
output: html_notebook
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

### Cost Burden 


```{r}
library(data.table)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(skimr)

fall18 <-fread("MCBSPUF18/puf2018_1_fall.csv")


fall18_reformatted <-fread("MCBSPUF18/PUF_1_FALL_reformatted.csv")


#ageclass=ADM_H_MEDSTA, insurance_type: Adm_op,dual etc

colnames(fall18)
```




```{r}

# sum(is.na(fall18))
dim(fall18)
# na.omit(fall18)
# sum(is.na(fall18))
fall18<-fall18[,c(1,4:238)]

dim(fall18)
fall18

colnames(fall18)
head(fall18)

fall18_reformatted  <- fall18_reformatted  %>%
 select(-SURVEYYR, -VERSION, -PUF_ID)
```


### Preprocessing

Data Dictionary:

ADM_H_MEDSTA: 1: elderly or 2: disabled or  3: unknown
ADM_H_GHPSW: 1: anyMA 0: no
ADM_H_PDRS: retiree drug subsidy(RDS) 1:not subsidized, 2: sub part year 3: sub full year, .:N/A, missing

insurance supplement type hierarchy being "Medicaid">"MAdv">"FFSwESI">"FFSwSelf">"FFSnoSupp"*/


```{r}
#Insurance Type
fall18<-fall18%>%mutate(ins_type=
                         case_when(
                            ADM_DUAL_FLAG_YR%in%c(2,3) ~ "Medicaid" ,
                            INS_D_MADV==1 ~"MAdv",
                            INS_D_PVESI==1&ADM_FFS_FLAG_YR%in%c(2,3)~"FFSwESI",
                            INS_D_PVSELF==1&ADM_FFS_FLAG_YR%in%c(2,3)~"FFSwSelf",
                        ADM_FFS_FLAG_YR%in%c(2,3)&INS_D_MADV==2&INS_D_PVESI==2&INS_D_PVSELF==2~"OnlyFFS",
                          TRUE ~"Missing"))


```


Demograhpics:

DEM_AGE: 1: disabled, 2: 65-74 3: elderly >=75 

new col: 'ageclass' 1: disabled 2: elderly

DEM_SEX: 1:male 2: female

DEM_RACE: 1:non hisp white 2: non hisp black 3: hisp 4: other

DEM_EDU: 1: no hs 2: hs / vocational/tech/business
3: more than hs

DEM_MARSTA: 1: married 2: widowed 3: divorced/separated 4:never married

new col 'marital_status': 1: married 2: widowed 3: single
DEM_INCOME: 1: <25k, 2:>=25k

DEM_CBSA: 1: metro 2: non-metro

DEM_IPR_IND: income poverty ratio medicare threshold 
1: <=100% of the Federal Poverty Level                            2  2: >100% and <=120% of the Federal Poverty                    3: >120% and <=135% of the Federal Poverty                        4: >135% and <=200% of the Federal Poverty                       5: >200% of the Federal Poverty Level 

ACC_HCDELAY: 1: yes 2:no
```{r}

#Creating new columns Ageclass and marital status
fall18<-fall18%>%
  mutate(ageclass=ifelse(fall18$DEM_AGE==1,1,2))%>%
  mutate(
    marital_status=ifelse(DEM_MARSTA==1,1,
                          ifelse(DEM_MARSTA==2,2,
                                 ifelse(DEM_MARSTA%in%c(3,4),3,0)
                                 )
                          )
        )


lapply(fall18[,-1], unique)



#Clean non response
fall18<-fall18 %>%
      mutate_all(~ replace(.,.%in%c("D","R",""), 99))


dim(fall18)


lapply(fall18,typeof)

#change char to num
fall18<-fall18%>%mutate_if(is.character,as.numeric)



```

# flag for checking whether insurance was not accepted by doc

Variables of significance:
MA_MADVYRS- those who have MCD for longer, do they have more costs?

```{r}






```


### Exploratory Data Analysis 

```{r}





```

